---
title: "Basic RLumUnmix functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic RLumUnmix functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(RLumUnmix)
library(Luminescence)
library(glue)
library(hash)

data_path <- system.file("extdata", package="RLumUnmix")

```

# Read/Create sequence
## Read .SEQ sequence
``` {r}
example_sequence <- Get_sequence(file_path = paste0(data_path, "/SEQ_file.SEQ"), lab_dose_rate = 0.2)
example_sequence


```

## Create sequence manually

```{r}
manual_sequence <- Create_sequence_manually()
manual_sequence

```

### Save as .RData

```{r}
saveRDS(manual_sequence, file = paste0(data_path, "manual_sequence.RData"), compress = FALSE)
```

### Now it can be read with Get_sequence()

```{r}
example_manual_sequence <- Get_sequence(file_path = paste0(data_path, "manual_sequence.RData"), lab_dose_rate = 0.2)
example_manual_sequence

```
# Compute OSL and TL metrics from synthetic/lab signals

## Generate synthetic OSL and TL signals
```{r}
data <- Generate_synthetic_OSL_TL_signals(sequence = example_sequence, model = "Bailey2001")
data

```
## Extract metrics from OSL and TL signals

```{r}
OSL_metrics <- Compute_OSL_metrics(sequence = example_sequence, data = Luminescence::get_RLum(data, recordType ='OSL$' , drop = FALSE))
TL_metrics <- Compute_TL_metrics(sequence = example_sequence, data = Luminescence::get_RLum(data, recordType ='TL$' , drop = FALSE), automatic_peak_finder =  TRUE)

```



### Show results for first TL curve metrics
```{r}
TL_metrics[['Record_1']]
```
## Extract metrics from signals stemming from .binx file
### Read binx file and extract metrics (only for sample located at position 1 in a standard risø TL/OSL reader)

Note that the automatic peak detection is not as good as for the synthetic signals. Argument "automatic_peak_finder" in function [Compute_TL_metrics()] can be set to FALSE to select peaks manually.

```{r}
bin_file <- read_BIN2R(file = paste0(data_path, "/binx_file.binx"), fastForward = FALSE)

print(glue("Position numbers: {unique(as.data.frame(bin_file)$POSITION)}"))

OSL_metrics <- hash()
TL_metrics <- hash()

sample_position_on_disk <- 1

    if ('OSL' %in% unique(data.frame(bin_file)[, "LTYPE"])) {
        OSL_data <- Risoe.BINfileData2RLum.Analysis(bin_file, pos = sample_position_on_disk, ltype = "OSL")
        OSL_metrics[sample_position_on_disk] <- Compute_OSL_metrics(sequence = example_sequence, data = OSL_data)
    }

    if ('TL' %in% unique(data.frame(bin_file)[, "LTYPE"])) {
        TL_data <- Risoe.BINfileData2RLum.Analysis(bin_file, pos = sample_position_on_disk, ltype = "TL")
        TL_metrics[sample_position_on_disk] <- Compute_TL_metrics(sequence = example_sequence, 
                                                                    data = TL_data, 
                                                                    automatic_peak_finder = TRUE)
    }


```

